---
title: "An introduction to heemod, a simple 3-state model"
output: 
  word_document:
    reference_docx: TEMPLATE.docx
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```



```{r user-functions}
dbloader <- function(packs) {
  system.time(
    if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
      print("you need the following packages to run this script")
      print(setdiff(packs, rownames(installed.packages())))
      print("Installing them now...")
      install.packages(setdiff(packs, rownames(installed.packages())))
      print("Now loading libraries...")
      sapply(packs, require, character.only = TRUE)
    } else {
      print("All packages are installed already")
      print("Loading the specified libraries...")
      sapply(packs, require, character.only = TRUE)
    }
  )
}

dbloader(c("heemod","simmer","tidyverse"))
```

# Introduction

This document is designed to go through the steps of setting up a simple 3 state partitioned survival CE model, using the heemod package. There are 3 primary aspects to consider. These are:  

1. Survival in each state
2. Cost of being in each state
3. Utility associated with each state

However, there are also other factors which can complicate the capturing of these 3 factors. Some examples include:

 - Costs changing over time within a state (e.g. Treatment duration limits, dosing schedules)
 - Time dependent utilities
 - Efficacy of an intervention depending on time since baseline
 - Stopping rules

In this first part, we will focus entirely on a very simple case, a 3-state model with constant hazards (i.e. exponential assumption), meaning that the transitions don't change over time.  

You will see that setting up the parameters sheet in this way, and within an rmarkdown document, has some advantages. It allows us to explain/annotate every parameter in the model one by one all in the same place. We can also use the text areas to link to bits of the code, so that the numbers we reference in the substantive text are linked to the data and results!

```{r Parameters sheet, echo=TRUE, message=FALSE, warning=FALSE}
#One of the best things about heemod is the ability to define an object containing all the parameters in the model, which then makes it much easier to keep track of the information being passed through the model

parameters_sheet <- define_parameters(
  
  # * control arm probabilities ------------------------------
  # from PFS
  t11_con = 0.9,
  t12_con = 0.05,
  t13_con = 0.05,
  #From PPS
  t21_con = 0,
  t22_con = 0.8,
  t23_con = 0.2,
  #from Dead
  t31_con = 0,
  t32_con = 0,
  t33_con = 1,
  
  # * Intervention arm probabilities ------------------------------
  # from PFS (This hypothetical intervention simply slows progression and reduces pre-progression death!)
  t11_int = 0.95,
  t12_int = 0.025,
  t13_int = 0.025,
  #From PPS
  t21_int = 0,
  t22_int = 0.8,
  t23_int = 0.2,
  #from Dead
  t31_int = 0,
  t32_int = 0,
  t33_int = 1,
  
  # * Current treatement flags  -------------------------------------------
  # This can be used to apply logic for which treatment the patient is currently on.
  #   This can help with applying subsequent treatment logic
  arm_cont = "On CONTROL treatment",
  arm_int  = "On INTERVENTION treatment",
  
  # * Costs      -------------------------------------------
  # Control drug is for life, one dose per cycle costing 50
  DrugCost_cont = 50,
  # New drug is given for maximum of 20 cycles, and only every 2 weeks, costing 150 each time:
  #   This part is very analogous to excel, ifelse works in exactly the same way!
  DrugCost_int  = ifelse(
    test = markov_cycle <= 20,
    yes = ifelse(
      test = markov_cycle %% 2 == 0,
      yes = 150,
      no = 0),
    no = 0
    ),
  #Resource use costs in each cycle, per cycle
  RU_PFS  = 10,
  RU_PPS  = 30,
  RU_Dead = 0,
  
  # * Utilities  -------------------------------------------
  #some standard state-based utilities
  U_PFS  = 0.95,
  U_PPS  = 0.55,
  U_Dead = 0
  
)


```

You can call a value from the parameters sheet as an "expression". This allows the package to treat anything as potentially a ball of logic, or just a number (similarly to excel). For instacne, the comparator cycle cost of £`r parameters_sheet$DrugCost_cont$expr` per cycle can be pulled form the code above this piece of text (See the source code: report.Rmd). Potentially, the parameters sheet can be derived from an external file, using the openxlsx package (which is much faster than the XLConnect package, due to the lack of a java dependency). This means that a separate workbook containing the basic (usually CiC) input information in excel can be kept safe.