---
title: "An introduction to heemod, a simple 3-state model. Session 1"
output:
  word_document:
    reference_docx: TEMPLATE.docx
  html_document:
    df_print: paged
  pdf_document: 
fontsize: "10pt"
geometry: "left=0.5in,right=0.5in"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```



```{r user-functions}
rm(list=ls())
dbloader <- function(packs) {
  system.time(
    if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
      print("you need the following packages to run this script")
      print(setdiff(packs, rownames(installed.packages())))
      print("Installing them now...")
      install.packages(setdiff(packs, rownames(installed.packages())))
      print("Now loading libraries...")
      sapply(packs, require, character.only = TRUE)
    } else {
      print("All packages are installed already")
      print("Loading the specified libraries...")
      sapply(packs, require, character.only = TRUE)
    }
  )
}

dbloader(c("heemod","simmer","tidyverse"))
```

# Introduction

This document is designed to go through the steps of setting up a simple 3 state partitioned survival CE model, using the heemod package. There are 3 primary aspects to consider. These are:  

1. Survival in each state
2. Cost of being in each state
3. Utility associated with each state

However, there are also other factors which can complicate the capturing of these 3 factors. Some examples include:

 - Costs changing over time within a state (e.g. Treatment duration limits, dosing schedules)
 - Time dependent utilities
 - Efficacy of an intervention depending on time since baseline
 - Stopping rules

In this first part, we will focus entirely on a very simple case, a 3-state model with constant hazards (i.e. exponential assumption), meaning that the transitions don't change over time.  

# Model 
## Parameters

You will see that setting up the parameters sheet in this way, and within an rmarkdown document, has some advantages. It allows us to explain/annotate every parameter in the model one by one all in the same place. We can also use the text areas to link to bits of the code, so that the numbers we reference in the substantive text are linked to the data and results!

```{r Parameters sheet, echo=TRUE, message=FALSE, warning=FALSE}
#One of the best things about heemod is the ability to define an object containing all the parameters in the model, which then makes it much easier to keep track of the information being passed through the model
# Another thing to note here is that this information could be extracted from an excel book using the openslsx package, which can extract directly from named ranges!
parameters_sheet <- define_parameters(
  
  # * functional parameters ----------------------------------
  #for any factors which need to be linked to cycle (1-month cycle length)
  year = floor(markov_cycle/12),
  
  # * demographics -------------------------------------------
  # Just age and sex
  age_start   = 60,
  age         = age_start + year,
  prop_female = 0.5,
  
  # * control arm probabilities ------------------------------
  #TODO: Make these time dependant - can they be linked to markov_cycle to allow for MSM?
  #TODO: confirm how to do this: I believe that the results from flexsurvreg can be used in the same way
  #       That we would use them in excel. just link them to markov_cycle (be careful of time unit, recommend weeks!)
  # from PFS
  t11_con = 0.9,
  t12_con = 0.05,
  t13_con = 0.05,
  #From PPS
  t21_con = 0,
  t22_con = 0.8,
  t23_con = 0.2,
  #from Dead
  t31_con = 0,
  t32_con = 0,
  t33_con = 1,
  
  # * Intervention arm probabilities ------------------------------
  # from PFS (This hypothetical intervention simply slows progression and reduces pre-progression death!)
  t11_int = 0.95,
  t12_int = 0.025,
  t13_int = 0.025,
  #From PPS
  t21_int = 0,
  t22_int = 0.8,
  t23_int = 0.2,
  #from Dead
  t31_int = 0,
  t32_int = 0,
  t33_int = 1,
  
  # * Current treatement flags  -------------------------------------------
  # This can be used to apply logic for which treatment the patient is currently on.
  #   This can help with applying subsequent treatment logic
  arm_cont = "On CONTROL treatment",
  arm_int  = "On INTERVENTION treatment",
  
  # * Costs      -------------------------------------------
  # Control drug is for life, one dose per cycle costing 50
  DrugCost_cont = 50,
  # New drug is given for maximum of 20 cycles, and only every 2 weeks, costing 150 each time:
  #   This part is very analogous to excel, ifelse works in exactly the same way!
  DrugCost_int  = ifelse(
    test = markov_cycle <= 20,
    yes = ifelse(
      test = markov_cycle %% 2 == 0,
      yes = 150,
      no = 0),
    no = 0
    ),
  #Resource use costs in each cycle, per cycle
  RU_PFS  = 10,
  RU_PPS  = 30,
  RU_Dead = 0,
  
  # * Utilities  -------------------------------------------
  #some standard state-based utilities
  U_PFS  = 0.95,
  U_PPS  = 0.55,
  U_Dead = 0
  
)


```

You can call a value from the parameters sheet as an "expression". This allows the package to treat anything as potentially a ball of logic, or just a number (similarly to excel). For instacne, the comparator cycle cost of £`r parameters_sheet$DrugCost_cont$expr` per cycle can be pulled form the code above this piece of text (See the source code: report.Rmd). Potentially, the parameters sheet can be derived from an external file, using the openxlsx package (which is much faster than the XLConnect package, due to the lack of a java dependency). This means that a separate workbook containing the basic (usually CiC) underlying input information in excel can be kept safe. It also makes redacting a model trivial!  

## Markov state transitions

Now that we have our parameters sheet we can begin to emplement the model. The underlying engine is the disease model, which follows a Markovian or semi-Markovian design, depending on how complicated the disease context is. In this case we are taking the simplest approach of transition probabilities which do not change over time. However, even if a more complicated logic surrounds the transistions, the below should remain the same. For instance, if t11_con (The control arm transition to go from state_A to stata_A, i.e. stay in the same state) were to follow some logic, that logic would be applied in parameters_sheet above, through linking the value to the variable markov_cycle (e.g. ifelse(markov_cycle<3,0.9,0.8) would allow the transition probability to change in cycle 3). Note that there will be an error if the probabilities do not sum to 1 as they should.  

```{r Transition matrices}
#define the transition probability matrix for the comparator
trans_cont <- define_transition(
  state_names = c("state_A","state_B","state_C"),
  t11_con, t12_con, t13_con,
  t21_con, t22_con, t23_con,
  t31_con, t32_con, t33_con
)

#define the transition probability matrix for the intervention
trans_int <- define_transition(
  state_names = c("state_A","state_B","state_C"),
  t11_int, t12_int, t13_int,
  t21_int, t22_int, t23_int,
  t31_int, t32_int, t33_int
)

trans_cont

trans_int

```

Now that the transition probabilities are defined, we can move onto defining the healthstates.

## Health state definitions

Here we can define the healthstates, which can contain logic, or can contain logic passed on from the parameters sheet (see *DrugCost_int*):

 - State A (PFS) has resource use costs only, and utility associated with PFS (drug costs added after)
 - State B (PPS) has resource use costs only, and utility associated with PPS
 - State C (Dead) has 0 cost and 0 utility

Obviously, this is a very simple set up. However, logic can be applied in here to allow for time dependencies etc (using markov_cycle)

## Defining the treatment strategies
Now that the transitions are defined, we need to define what the alternative treatment strategies are. This is done using define_strategy. Logic can also be applied here, but please keep it to the treatment-arm specific factors. This way, it is clear that the below code is only to set the treatment strategies!


```{r defining treatment arms}
#control
control_arm <- define_strategy(
  transition = trans_cont,
  state_A = define_state(
    cost    = DrugCost_cont + RU_PFS,
    utility = U_PFS
  ),
  state_B = define_state(
    cost    = RU_PPS,
    utility = U_PPS
  ),
  state_C = define_state(
    cost = 0,
    utility = 0
  )
)
#intervention
intervention_arm <- define_strategy(
  transition = trans_int,
  state_A = define_state(
    cost    = DrugCost_int + RU_PFS,
    utility = U_PFS
  ),
  state_B = define_state(
    cost    = RU_PPS,
    utility = U_PPS
  ),
  state_C = define_state(
    cost = 0,
    utility = 0
  )
)

```


## Running the model

That's it! we have set up the transitions, and the states. Now all that is left is to run the model. We need to give it the parameters sheet, the strategies, the number of cycles etc.

```{r Running the model}
MODEL <- run_model(
  parameters = parameters_sheet,
  init = c(1,0,0),
  control_arm,
  intervention_arm,
  cycles = 200,
  cost = cost,
  effect = utility
)
```

To take the markov trace and turn it into a partitioned survival graph:

```{r patient flow}
patient_flow <- get_counts(MODEL)
patient_flow[which(patient_flow$state_names == "state_C"),"count"] <- 1-patient_flow[which(patient_flow$state_names == "state_C"),"count"]
patient_flow <- filter(patient_flow,patient_flow$state_names != "state_B")

ggplot(patient_flow, aes(x = markov_cycle, y = count, colour = `.strategy_names`)) + 
  geom_line(aes(linetype = state_names)) + 
  theme_classic() +
  theme(legend.position = "bottom") + 
  scale_x_continuous(expand = c(0,0)) + 
  scale_y_continuous(expand = c(0,0))

```

Or alternatively, a standard markov trace can be produced:
```{r}
plot(MODEL) + 
  theme_classic() +
  theme(legend.position = "bottom") + 
  scale_x_continuous(expand = c(0,0)) + 
  scale_y_continuous(expand = c(0,0))
```



general results:

```{r model summary}
summary(MODEL)
```

